---
title: "Assignment 2"
author: "Tor Jensen, Timo Klingler, Margaret Koehn and Janette Wiget"
date: "8 November 2015"
output:
  html_document: default
  pdf_document:
    latex_engine: xelatex
---

```{r load_libraries, include = FALSE}
#loading the required packages for the report and set options
require("rvest")
require("xml2")
require("readr")
require("stringr")
require("ggplot2")
require("ggthemes")
require("scales")
require("viridis")
require("raster")
require("sp")
require("splitstackshape")
require("plyr")
require("dplyr")
options(scipen=999)
```

The following data analyis is based on data from the website "I paid a pribe" [link](http://www.ipaidabribe.com) and addresses the topic of corruption. Especially in the developing countries, as india in our case, we can observe a high occurence of corruption. Corruption has many dimensions which all are related to political, economic, social and environmental effects. Because all of these effects have a serious impact on the individual communities and countries, it is essential to take the aspect of corruption into account when analysing a certain economy. Our data analysis will provide general information about the occurrence of corruption in india and will give an overview regarding its distrubtion along certain subsets:

  * Dataset about the last 1000 reports
      + Scraping data from web
      + Data cleaning
  * Data analysis
      + Basic summary of dataset
      + Records per day
      + Relationship between views and amount
      + Getting the right understanding of the numbers
      + GDP per Capita and Amount of bribes paid
      + Total Amount of Bribe: Geographical plot
      
---

## Creating the dataset about the last 1000 reports

```{r scrape_data, eval = FALSE, include = FALSE}
# Creating a dataset by scraping data from the "I Paid a Bribe"" website, linking to pre-scraped dataset below this block

# General considerations about the layout of the website:  
# Data about the bribe; time, place, amount, category, etc. is available from the main search page.  
# The full text describing the event is only available from the link itself. 
# Can get most of the data from the main "search", without having to go through every link that has been pulled

## Initialize dataframe with the desired data columns
df <- data_frame(title = character(), 
                 amount = character(), 
                 dep = character(), 
                 trans = character(), 
                 views = character(),
                 city = character(),
                 date = character(), 
                 time = character())

## initialize empty dataframe for 
dftemp <- data_frame()

## Scrape data from "I paid a bribe" website

for (i in 0:99) { #loop through first 100 pages, 10 results per page = 1000 
  link <- paste("http://www.ipaidabribe.com/reports/paid?page=",i*10, sep = "") # Create hyperlink based on loop function
  print(paste("processing", i, sep = " ")) # progress report
  main <- read_html(link, encoding = "UTF-8") # define the static part of link references
  
  title <- main %>% 
    html_nodes(".heading-3 a") %>% 
    html_text() 
  
  amount <- main %>% # feed `main.page` to the next step
    html_nodes(".paid-amount span") %>% # get the CSS nodes
    html_text() # extract the link text
  
  dep <- main %>% # feed `main.page` to the next step
    html_nodes(".name a") %>% # get the CSS nodes
    html_text() # extract the link text
  
  trans <- main %>% # feed `main.page` to the next step
    html_nodes(".transaction a") %>% # get the CSS nodes
    html_text() # extract the link text
  
  views <- main %>% # feed `main.page` to the next step
    html_nodes(".overview .views") %>% # get the CSS nodes
    html_text() # extract the link text
  
  city <- main %>% # feed `main.page` to the next step
    html_nodes(".location") %>% # get the CSS nodes
    html_text() # extract the link text
  
  date <- main %>% # feed `main.page` to the next step
    html_nodes(".date") %>% # get the CSS nodes
    html_text() # extract the link text
  
  time <- main %>% # feed `main.page` to the next step
    html_nodes(".time-span") %>% # get the CSS nodes
    html_text() # extract the link text
  
  dftemp <- cbind(title, amount, dep, trans, views, city, date, time) # bind the variables together into a 10 by n dataframe
  df <- rbind(df,dftemp) # rbind the temp dataframe for this page to the main dataframe
  
  Sys.sleep(1) # timer, wait 1 second
  cat(" done!\n") #progress report
}

#clean unused variables from workspace
rm("title", "amount", "dep", "trans", "views", "city", "date", "time", "dftemp", "i")

## Split the city column into a city and state column

df$states <- lapply(strsplit(as.character(df$city), "\\,"), "[", 2)
df$city <- lapply(strsplit(as.character(df$city), "\\,"), "[", 1)
```

```{r clean_data, echo = FALSE}
## Clean and order dataset, force correct data types, factors for faceting and checking number of levels
clean <- function(df) {
df$title <- df$title %>% # clean text
  str_replace_all(pattern = "\\n" , replacement = " ") %>%
  str_trim()

df$dep <- df$dep %>% 
  as.factor() # convert to factor 

df$trans <- df$trans %>% 
  as.factor() # convert to factor 

df$amount <- df$amount %>% # clean text from amount and convert to numeric
  str_replace_all(pattern = "Paid INR" , replacement = " ") %>% 
  str_replace_all(pattern = "," , replacement = "") %>% 
  str_trim() %>% 
  as.numeric() 

df$views <- df$views %>% # clean text from views and convert to numeric
  str_replace_all(pattern = "views" , replacement = " ") %>%
  str_trim() %>% 
  as.numeric()

df$city <- df$city %>% # clean text from city
  as.character() %>%
  str_trim() %>%
  as.factor() # conver tot factor

df$states <- df$states %>% # clean text from states
  as.character(df$states) %>%
  str_trim() %>% 
  as.factor() # convert to factor

df$time <- as.numeric(str_extract(df$time,"[0-9]*"))*!grepl("minutes|hours",df$time) # clean hours and minutes out of time stamp and change to whole number of days

df$date <- as.Date(df$date, format("%B %d, %Y")) # convert the date column to date format

df <- df[, c(1,2,3,4,5,6,9,7,8)] # rearrange columns

return (df)

}
```

After having applied string manipulation and other data cleaning methods we are able to have a first look at the values by ordering the number. First, we observed a single value that is incredibly high. The amount of money is equivalent to 113 million euros - and this "only" for a birth certificate. This report seems highly unlikely and was removed because of the distortion it would've had introduced in the analysis. Second, the head function reveals quite a few instances of Rs 1. Due to the poverty of a large portion of the Indian population, we cannott really argue for setting a minimum threshold for bribes even at this relatively low amount. 


```{r cleaning_errors_outliers, eval = FALSE, include = FALSE}
# futher Data cleaning

## look at headers and footers for obvious outliers in bribe size

### sorted in descending value 
df %>% 
  arrange(desc(amount)) %>% 
  head(5) 

### sorted in ascending order
df %>% 
  arrange(desc(desc(amount))) %>% 
  head(5)

## look for missing values in the character variables (title, dep, trans, city, state)
dfmissing <- df %>% filter(title == "" | dep == "" | trans == "" | city == "" | states == "") %>% head(5)
# only a single entry with no data

## count number of rows with NA values
narows <- nrow(df[!complete.cases(df),])
# total number is 0, no need to filter out NA values

## test for misassigned values in either 
duplicate <- match(df$city,df$states)
duplicate2 <- match(df$states,df$city)
dupcheck <- df[!is.na(duplicate),]
dupcheck2 <- df[!is.na(duplicate2),]
# 3 values come up, city name is capital name, chandigarh is a state and an area so no problem

## remove unwanted variables again used
rm("dfmissing", "narows", "duplicate", "duplicate2", "dupcheck", "dupcheck2")

```

```{r import_csv, include = FALSE}
# Import the common csv file from github to work on, at end of project remove this reference, lines to be removed marked with [X]

df_raw <- read_csv("https://raw.githubusercontent.com/TorJensen/Assignment-2/master/df.csv") # read csv file from github [X]
df_raw <- df_raw[,2:10] # remove the rownames column - [X]

df <- df_raw # define df
rm(df_raw) # remove df_raw from the environment
```

---

## Basic summary of dataset

```{r summary, echo = FALSE}
df <- clean(df) #apply cleaning function
df <- df %>% 
  filter(dep != "", amount < max(amount)) # Remove entry with missing data and remove the maximum value

# exploratory data analysis 

summary(df %>% select(title, amount, views, states, date))
```

The basic summary shows large differences between the mean and median amounts and the third quartile. The mean difference indicates that a relatively small amount of large bribes are increasing the mean, while the vast majority of bribes reported are relatively small. Almost half of the bribes are reported from Bangalore, and about 2/3rds of the bribes are recorded within the top 5 departments. The reporting period ends on the 12th of October, so we only have 1 month worth of data, which limits the time/weekeday/seasonality analysis options available. 

Having a closer look at the distribution of bribes submitted per day, we can see that the majority of information in our dataset was submitted on one single day in the last month (see below). This is definitely an outlier compared to the other days wheer we count about 10-30 reports on average per day. Having a look at the datasets' other variables assigned to that specific day, it seems like that all the other values of the varibles did not show any significant variances and thus can be said that the bribes on that day are reasonably well-distributed across states and departments. The best possible explanation connected to this observation is probably the fact of an excisting barrage of spam or simply incorrect reports.  

```{r records_per_day, echo = FALSE}
# plot showing amount of records per day
p <- ggplot(df %>% group_by(date) %>% summarise(count = n()),aes(date, count))
p + geom_bar(stat = "identity") + 
  labs(x="Date", y="Count", title="Records per day") +
  theme(plot.title = element_text(lineheight=.8, face="bold", vjust=1))
```

Although the chance that we are dealing with an error source here is quite likely, a quick google news search for october reveals that an article was posted in Times of India on the 11th of October which could still be a possible explanation for the sudden spike or at least for a part of it [link](http://timesofindia.indiatimes.com/city/chennai/Chennai-corporation-takes-no-action-on-corruption/articleshow/49316078.cms). The article is not only reporting on corruption, but also is directly refering on the "I paid a bribe" website as the have gathered information from the webpage.

```{r narrow_data, echo = FALSE}
# narrow down the dataset to entries made on October 12
df_lump <- df %>% 
  filter(date == min(date)) # filter the data based on the earliest day in the dataset

summary(select(df_lump, amount, dep, trans, views, city, states)) # show summary data
```

According to the previous summary there does not seem to be any significant relation. We can only observe few outliers which are for medium-sized amounts. Consequently, we can assume that there must be other intervening factors that have a larger impact on these outliers other than size alone.

---

## Relationship between views and amount paid

Building on the previous discussion, another possible relation to have a further look at is the amount of views compared to the amount of money paid. As shown in the graph below, it is difficult to argue for a strong relation between those two factors. 

```{r cluster_data, echo = FALSE}
# looking at the relation between views and amount of data - do higher value bribe reports attract more attention?
p2 <- ggplot(df,aes(views, log(amount)))
p2 + geom_point(alpha = 0.4) + # lower alpha (transparency) to be able to identify clusters of data
  labs(x="Views", y="log(Amount)", title="Relationship Between Views and Amount of Data") +
  theme(plot.title = element_text(lineheight=.8, face="bold", vjust=1))
```

---

## Getting the right perspective about the numbers

There are 27 states and 17 departments (with 40 transaction types) in the 1000 records we have mined, so an analysis looking at transaction type level by state, for example, would need more than 1000 data points in order to be able to get a reasonable analysis (assuming some degree of normality.) Instead we decided to focus on higher level summaries, specific departments or specific states. 

```{r levels, echo = FALSE}
levelcount <- llply(df,levels) %>% summary()
levelcount <- ldply(llply(llply(levelcount[,1],str_trim),as.numeric)) %>% filter(V1>0)
names(levelcount) <-c("Variable","Factors")
levelcount
```

In the gathered data we can see that most bribes are paid to the department of Municipal Services, especially for issuing a birth certification. Facing this situation, we wanted to find out if there are any reasons for the high rate of corruption in this sector. Apart from the fact that the Indian citizens mostly have to pay bribes for every service they ask for, the high birth rate in India is probably the main reason for this circumstance. The more urgent the need for a document, the higher the amount the people have to pay for a certain service. Furthermore, we have aggregated the amounts of bribes paid per department. The resulting visualisations are showing us that there is not just a wide selection of corruption in the sector of Municipal Services, but also generate these public authorities by far the highest amount of money. Relating to the high birth rate in India, as mentioned above, it is probably the easiest and most common way to be bribed. Due to the complicated tax and licensing systems and corrupt police system, the bribe payments in the departments of “Income Tax” and “Police” belong to the highest in India.

```{r bribes_sliced, echo = FALSE}
# Number of bribes and their sizes, by variables in various combinations, these can be called throughout the text:

### below are some tables looking at the... 

#...department
df_dep <- df %>% 
  group_by(dep) %>% 
  summarise(count = n(), amount = sum(amount))%>% 
  mutate(avgamount = round(amount/count)) %>% 
  arrange(desc(avgamount))

#...state
df_state <- df %>% 
  group_by(states) %>% 
  summarise(count = n(), amount = sum(amount))%>% 
  mutate(avgamount = round(amount/count)) %>% 
  arrange(desc(count))

#...department and transaction type
df_deptrans <- df %>% 
  group_by(dep,trans) %>% 
  summarise(count = n(), sum = sum(amount))%>% 
  mutate(avgamount = round(sum/count)) %>% 
  select(dep, trans, count, avgamount)


df_deptrans[with(df_deptrans, order(desc(count))),]
#...state and city
df_statecity <- df %>% 
  group_by(states, city) %>% 
  summarise(count = n(), sum = sum(amount)) %>% 
  mutate(avgamount = round(sum/count))

#...state and department, the thougth was that some departments could be more corrupt in some areas compared to others, hard to tell with so few data for so many sates. 
df_statedep <- df %>% 
  group_by(states, dep) %>% 
  summarise(count = n(), sum = sum(amount)) %>% 
  mutate(avgamount = round(sum/count)) %>% 
  arrange(desc(count))

#...number of views by state, are some states viewed more frequently, do states with lots of reports have more views, do states with a higher value of bribes have more views
df_stateviews <- df %>% 
  group_by(states) %>% 
  summarise(count = n(), views = sum(views), amount = sum(amount)) %>% 
  mutate(avgviews = round(views/count), avgamount = round(amount/count))

#...number of views by department, are some departments viewed more frequently, do departments with lots of reports have more views, do departments with higher bribe value have more views.  
df_depviews <- df %>% 
  group_by(dep) %>% 
  summarise(count = n(), views = sum(views), amount = sum(amount)) %>% 
  mutate(avgviews = round(views/count), avgamount = round(amount/count))

#Does number of views level out with time? 
df_viewsdate <- df %>% 
  group_by(date) %>% 
  summarise(count = n(), views = sum(views)) %>% 
  mutate(avgviews = round(views/count))

head(df_deptrans[with(df_deptrans, order(desc(count))),],5) 
head(df_deptrans[with(df_deptrans, order(desc(avgamount))),],5) 
```

We can observe that the most pervasive bribes aren't necessarily the largest one. Although this isn't very suprising, these graphs and tables emphasize the risk that you possibly have to face corruption everywhere you go and  in every interaction the state is involved. Ration cards are especially important as they are necessary to benefit from the fuel subsidies the indian government gives, and to access the public food distribution systems. For the less frequent but higher value amounts, it is striking that the issuing of [new PAN cards](http://www.indiacgny.org/pdf/pan_faq.pdf), which are not just used to pay taxes, but also as proof of identity and are necessary to enter most financial transactions, are either bought on false from corrupt officials, or those officials extort people who need it. 

The graphs below display the top 5 departments ordered by the number of bribes reported and the transactions with the highest amounts in average.

```{r bars, echo = FALSE}

### Bribes per department

bribe_dep <- aggregate(amount ~ dep, df, sum)
bribe_dep$dep <- as.character(bribe_dep$dep)

bribe_dep[which(with(bribe_dep, dep == "Municipal Services")), c(1)] <- "Mun. Services"
bribe_dep[which(with(bribe_dep, dep == "Food, Civil Supplies and Consumer Affairs")), c(1)] <- "Food/Civil Supplies"
bribe_dep[which(with(bribe_dep, dep == "Stamps and Registration")), c(1)] <- "Stamps and Reg."
bribe_dep[which(with(bribe_dep, dep == "Commercial Tax, Sales Tax, VAT")), c(1)] <- "Com./Sales/VAT"
bribe_dep[which(with(bribe_dep, dep == "Customs, Excise and Service Tax")), c(1)] <- "Cust./Ex./Serv"

bribe_dep <- bribe_dep %>% filter(dep != "") %>% arrange(desc(amount))

bribe_dep$dep <- as.factor(bribe_dep$dep)

bribe_dep$dep <- factor(bribe_dep$dep,levels(bribe_dep$dep)[c(8, 6, 11, 15, 2, 9, 16, 5, 3, 10, 14, 13, 4, 7, 17, 1, 12)])

p3 = ggplot(bribe_dep, aes(x = dep, y = amount, fill = dep))
p3 + geom_bar(stat="identity") + 
  theme_tufte() + 
  scale_y_continuous(labels=comma) +
  labs(x="", y="Amount", title="Bribe Amount per Department") +
  theme(axis.text.x = element_blank(),
        legend.position = "none", 
        plot.title = element_text(lineheight=.8, face="bold", vjust=1)) + 
  facet_wrap(~ dep, scales = "free_x") +
  scale_fill_viridis(discrete = TRUE)
```

---

## GDP per Capita and the amount of bribes paid

The following code lines and its resulting graph are questioning the relation between the GDP per capita of each state and the average amount of money paid per bribe in a certain state. In order to get the needed information, the data from our dataset is complemented by tables scraped from two different wikipedia sources.

```{r gdp_per_capita, echo = FALSE}
# Bribe amounts
bribe_sta <- aggregate(amount ~ states, df, sum)
bribe_n <- df %>% group_by(states) %>% count(states) # create column with number of bribes per state 
bribe_sta$n <- bribe_n$n # add column with number of bribes per state to amount of bribes per state
bribe_sta <- bribe_sta[-c(9),] # remove Hardoi
bribe_sta$states <- as.character(bribe_sta$states)
bribe_sta[which(with(bribe_sta, states == "Orissa")), c(1)] = "Odisha" #rename columns

# Download India GDP data

ind.gdp = read_html("https://en.wikipedia.org/wiki/List_of_Indian_states_by_GDP") %>%
  html_nodes(xpath = '//*[@id="mw-content-text"]/table[2]') %>% 
  html_table() %>% data.frame # then convert the HTML table into a data frame

names(ind.gdp) <- c("states", "co_ru")

ind.gdp$co_ru = as.numeric(gsub("," , "" , ind.gdp$co_ru)) # remove commas from indian number notation
ind.gdp = ind.gdp %>% mutate(gdp = co_ru * 10000000) %>% as.data.frame() %>% select(states, gdp)# 1 crore is 10 million, convert to unit measure and stick it in a new dataframe
ind.gdp[which(with(ind.gdp, states == "Chattisgarh")), c(1)] = "Chhattisgarh" 
ind.gdp[which(with(ind.gdp, states == "Jammu & Kashmir")), c(1)] = "Jammu and Kashmir" 
ind.gdp[which(with(ind.gdp, states == "Andaman & Nicobar Islands")), c(1)] = "Andaman and Nicobar Islands" 

# Download India population data

ind.pop = read_html("https://en.wikipedia.org/wiki/List_of_states_and_union_territories_of_India_by_population") %>%
  html_nodes(xpath = '//*[@id="mw-content-text"]/table[2]') %>% 
  html_table() %>% data.frame %>% select(State.or.union.territory, Population..2011.Census..12.....of.Population.of.India..13.)

names(ind.pop) <- c("states", "population")

ind.pop <- ind.pop[-c(37), ] # remove rows 37, India, from population
ind.pop[which(with(ind.pop, states == "Manipurβ")), c(1)] = "Manipur"
ind.pop$population <-  str_extract(ind.pop$population, "(?<=\\d{19}...)[0-9,]*") %>% str_replace_all(",","") %>% as.numeric()

# Combine bribe_sta, ind.gdp and ind.pop

ind.df <- rbind(bribe_sta, data.frame(states = setdiff(ind.pop$states, bribe_sta$states), amount = rep(0,length(setdiff(ind.pop$states, bribe_sta$states))), n = rep(0, length(setdiff(ind.pop$states, bribe_sta$states)))))

ind.gdp <- rbind(ind.gdp, data.frame(states = setdiff(ind.df$states, ind.gdp$states), gdp = rep(0,length(setdiff(ind.df$states, ind.gdp$states)))))

ind.df <- left_join(ind.df, ind.gdp, by = c("states" = "states"))
ind.df <- left_join(ind.df, ind.pop, by = c("states" = "states"))
ind.df <- mutate(ind.df, gdp_cap = gdp/as.numeric(population)) #create new colun with gdp per capita
ind.df <- mutate(ind.df, m_per_b = amount/n)

# Remove NA values from dataset
          
rm.nan <- function(x)
do.call(cbind, lapply(x, is.nan))

ind.df[rm.nan(ind.df)] <- 0
```

The resulting graph from above is raising the question whether states with a higher GDP per capita have a higher/lower frequency of people paying bribes. The question relating to a general relationship between those to factors can best be examined by the performance of a regression analysis.
According to the regression plot there does not seem to be a strong relaton between GDP per Capita and the average amount of money spent. Nevertheless, it can be argued that there is a slight tendancy of moving towards the fact that states with a higher GDP per Capita come along with a higher amount paid per bribe than average. Which would obviously also make sense as those people have more ressources to spend.

```{r gdp_bribe_graph, echo = FALSE}
p4 <- ggplot(ind.df, aes(gdp_cap, log(m_per_b)))
p4 + geom_point() + 
  labs(x="GDP per Capita", y="log(Amount per Bribe)", title="GDP/Capita and Amount of Bribes Paid") +
  theme(plot.title = element_text(lineheight=.8, face="bold", vjust=1))
```

In the last part of our data analysis we have created a map of India showing the total amount of bribes by each state. The resulting map illustrates, that the highest number of bribes can be observed in the Southern part of India, particularly in Kerala, Tamil Nadu, Karnataka. Jammu and Kashmir has also an high amount of bribes. In conclusion, we can say that the bribes are more concentrated in the North and South of India, near the borders to Pakistan and the Indian Ocean. Here we have a total amount of bribes paid of about 2,000,000.  In the inside of the country the level of bribes paid decreases in comparison to the external states. One reason for this circumstance could be the high number of citizens in these regions compared to the other states in India. We could also speculate about the question whether people have less internet access in the inner states of the countries or whether their level of education is lower so that they simply don't have the knowledge about the website we're taking the data from. 

```{r visualize_map, message = FALSE, echo = FALSE}

## Visualize bribe/state

ind.map <- getData("GADM", country="IND", level=1)

bribe_s <- select(bribe_sta, states, amount)

am.count <- rbind(bribe_s, data.frame(states = setdiff(ind.map$NAME_1, bribe_s$states), amount = rep(0, 12)))

am.count <- am.count[-c(38),]
am.count <- am.count[- grep("Andaman and Nicobar", am.count$states),]

am.count[which(with(am.count, states == "Uttarakhand")), c(1)] <- "Uttaranchal"

am.count <- am.count[order(am.count$states, decreasing = FALSE),]

ind.map <- fortify(ind.map)

am.count$index <- as.numeric(tapply(ind.map$id, ind.map$id, length))
am.count <- data.frame(amount = am.count$amount, index = am.count$index)

am.count <- expandRows(am.count, "index")

ind.map$count <- am.count$amount

p5 <- ggplot() + 
  geom_polygon(data = ind.map, aes(x=long, y=lat, group=group, fill = count), colour = "black") +
  scale_fill_gradient("Bribes paid \nby state ", low="white",high="black", breaks = c(0, 1e+06, 2e+06, 3e+06, 4e+06),
                      labels=c("0", "1M", "2M", "3M", "4M")) +
  labs(x="", y="", title="Total Amount of Bribes by State") + 
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank(), 
        axis.ticks.x = element_blank(),axis.text.x = element_blank(), 
        plot.title = element_text(lineheight=.8, face="bold", vjust=1))
p5
```
